---
title: "Exercise-08"
format: html
editor: visual
---

```{r}
# Download the file and do a quick summary
f <- "https://raw.githubusercontent.com/difiore/ada-datasets/main/Street_et_al_2017.csv"
library(readr)
d <- read_csv(f, col_names = TRUE)
library(skimr)
skim(d)
```

Plot ECV as a function of social group size (Group_size), longevity (Longevity), juvenile period length (Weaning), and reproductive lifespan (Repro_lifespan)

```{r}
library(ggplot2)
# ECV vs. Social Group Size
ggplot(d, aes(x = Group_size, y = ECV))+
  geom_smooth(formula = y ~ x, method = "lm")+
  labs(title = "Brain Size vs Social Group Size")

# ECV vs. Longevity 
ggplot(d, aes(x = Longevity, y = ECV))+
  geom_smooth(formula = y ~ x, method = "lm")+
  labs(title = "Longevity vs Social Group Size")

# ECV vs. Weaning
ggplot(d, aes(x = Weaning, y = ECV))+
  geom_smooth(formula = y ~ x, method = "lm")+
  labs(title = "Juvenile Period Length vs Social Group Size")

# ECV vs Reproductive Lifespan
ggplot(d, aes(x = Repro_lifespan, y = ECV))+
  geom_smooth(formula = y ~ x, method = "lm")+
  labs(title = "Reproductive Lifespan vs Social Group Size")

```

Calculating linear regression coefficients

```{r}
library(tidyr)
d <- d |> drop_na(ECV)
d <- d |> drop_na(Group_size)

# --- Calculate by hand ---
groupsize <- d$Group_size
ecv <- d$ECV

B1 <- sum((groupsize - mean(groupsize))* (ecv - mean(ecv))/sum(groupsize - mean(groupsize))^2)

B0 <- mean(ecv) - B1*mean(groupsize)

# --- Confirm with lm() function ---
model <- lm(ECV ~ Group_size, data = d)
summary(model)
```

Repeat the analysis

```{r}
# Calculate seperate regressions by group 
library(purrr)
models_cata <- lm(ECV ~ Group_size,
                  data = subset(d, Taxonomic_group == "Catarrhini"))

models_platy <- lm(ECV ~ Group_size,
                   data = subset(d, Taxonomic_group == "Platyrrhini"))

models_streps <- lm(ECV ~ Group_size,
                   data = subset(d, Taxonomic_group == "Strepsirhini"))

# confirm the coefficients
coef(models_cata)
coef(models_platy)
coef(models_streps)

# Fit one combined model with an interaction to double check
interaction_model <- lm(ECV ~ Group_size * Taxonomic_group, data = d)
interaction_model

# Plot the models to visualize it and tirple check
ggplot(d, aes(x = Group_size, y = ECV, color = Taxonomic_group)) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "ECV vs Group Size by Primate Radiation")
```

Answer: Yes, the regression coefficients differ among groups. To determine that, I calculated the individual coefficients of each group and visualized it to help me read the difference better.

Calculate the standard error for the slop coef, 95% CI, P-value by hand

```{r}
# The easy way to know
summary(model)

# Calculate by hand
n <- length(groupsize)

#slope
B1 <- sum((groupsize - mean(groupsize))* (ecv - mean(ecv))) / sum((groupsize - mean(groupsize))^2)

B0 <- mean(ecv) - B1*mean(groupsize)

#Residuals
residuals <- ecv - (B0 + B1*groupsize)

#Residual variance
sigma2 <- sum(residuals^2) / (n - 2)

# SE of slope
SE_B1 <- sqrt(sigma2 / sum(groupsize - mean(groupsize))^2)

# 95% CI
t_value <- qt(0.975, df = n - 2)
CI_lower <- B1 - t_value * SE_B1
CI_upper <- B1 + t_value * SE_B1

#P-value for slope
t_stat <- B1 / SE_B1
p_value <- 2 * (1 - pt(abs(t_stat), df = n -2))

summary(model)$coefficients
confint(model)
```

Permutation approach with 1000 permutations for the slope coefficient

```{r}
set.seed(123)
n_perm <- 1000
x <- d$Group_size
y <- d$ECV

#Original Slope
B1_obs <- coef(lm(y~x))[2]

B1_perm <- numeric(n_perm)

for(i in 1:n_perm) {
  y_perm <- sample(y) 
  B1_perm[i] <- coef(lm(y_perm ~ x))[2]
}

hist(B1_perm)

#calculate p-value: theory-based
B1_mean <- mean(B1_perm)
B1_sd   <- sd(B1_perm)

# z-score of observed slope
z <- (B1_obs - B1_mean) / B1_sd

# Two-sided p-value
p_theory <- 2 * (1 - pnorm(abs(z)))
p_theory
```

Bootstrapping to generate a 95% CI

```{r}
set.seed(123)
n_boot <- 1000
B1_boot <- numeric(n_boot)

x <- d$Group_size
y <- d$ECV
n <- nrow(d)

for(i in 1:n_boot){
  indices <- sample(1:n, size = n, replace = TRUE)
  x_boot <- x[indices]
  y_boot <- y[indices]
  
  model_boot <- lm(y_boot ~ x_boot)
  B1_boot[i] <- coef(model_boot)[2]
}

# Quantile-based 95% CI
CI_quantile <- quantile(B1_boot, probs = c(0.025, 0.975))
CI_quantile

# Theory-based 95% CI
B1_mean <- mean(B1_boot)
B1_sd <- sd(B1_boot)
CI_theory <- c(B1_mean - 1.96 * B1_sd, B1_mean + 1.96 * B1_sd)
CI_theory

# check if the slop is different from zero
B1_obs <- coef(lm(y ~ x))[2]
CI_quantile
CI_theory
```

Answer: Yes, both CIs suggest that my slop coefficient is different from zero.
